---
title: "R Notebook"
output: html_notebook
---

## About the data.

### Introduction.

This database contains information on tennis matches in a period from 2000 to 2016.

On this occasion, I will analyze the behavior of the players in which I will determine: The cities played, the players, the field conditions, and the years played.

In addition to that, the Igaming agencies will also be added with the predictions they had before the matches, to determine which is the best agency to go to bet.

#### Load libraries 

```{r}
library(dplyr)
library(RSQLite)
library(grid)  
library(gridExtra) 
library(ggExtra)
library(cowplot)
library(tidyverse)
library(ggvis)
library(shiny)
library(DT)
library(stringr)
library(knitr)
library(kableExtra)
library(remotes)
library(data.table)
library(ggplot2)
library(reshape2)
```

#### Read the data

```{r}
str(Data)
summary(Data)
```

#### Clean de data and prepare the data frames

```{r}
#Converting the Rank variables into numeric variables
Data$LRank=as.numeric(Data$LRank)
Data$LPts=as.numeric(Data$LPts)

#Formatting Date to extract Season Info
Data$Date=as.Date(Data$Date,format='%d/%m/%Y')
Data$Season=as.numeric(format(Data$Date,"%Y"))

Data %>% 
  mutate(Winner = gsub(" $", "", Winner), Loser = gsub(" $", "", Loser))
summary(Data)
```

### Let's see the cities

Within our analysis, we can see that our favorite cities are: Paris, London, New York, and Melbourne.

```{r}

fillColor = "#FFA07A"
fillColor2 = "#F1C40F"

Data %>%
  group_by(Location) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  ungroup() %>%
  mutate(Location = reorder(Location,Count)) %>%
  head(20) %>%
  
  ggplot(aes(x = Location,y = Count)) +
  geom_bar(stat='identity',colour="white", fill = fillColor) +
  geom_text(aes(x = Location, y = 1, label = paste0("(",Count,")",sep="")),
            hjust=0, vjust=.5, size = 4, colour = 'black',
            fontface = 'bold') +
  labs(x = 'City', 
       y = 'Count', 
       title = 'Top 20 cities played') +
  coord_flip() +
  theme_bw()
```

## Let's analyse the players.

```{r}
#Winners
winners = Data %>% 
  group_by(Surface, Winner) %>% 
  summary (nWin = n()) 


#Losers
losers= Data %>% 
  group_by(Surface, Loser) %>% 
  summary(nLose = n()) 

#Format to the new tables
colnames(winners)[2] = "Name"
colnames(losers)[2] = "Name"

```

Let's take de percent of winners against losers

```{r}

players_bySurface = merge(winners, losers, by = c("Surface", "Name"), all = T) %>% 
  mutate(nWin = ifelse(is.na(nWin), 0, nWin), nLose = ifelse(is.na(nLose), 0, nLose), winPerc = nWin/(nWin+nLose)) %>% 
  arrange(desc(winPerc))

#Some format of the percet of each player
players_overall = data.frame(players_bySurface %>% 
                               group_by(Name) %>% 
                               summarise(nWin = sum(nWin), nLose = sum(nLose)) %>% 
                               mutate(winPerc = paste(round(nWin/(nWin+nLose)*100,2),"%",sep="")) %>% 
                               filter(nWin+nLose>100) %>% 
                               arrange(desc(winPerc)))
```

Let's see our table, to have a better knowledge of what are we doing.

These tables are showing us the stats of all the players I am analyzing to get to know data such as Names, number of games won, number of games loosed, percentage of the total games won according to their games played.

```{r}
players_overall %>% 
  select(Name, nWin, nLose, winPerc) %>% 
  datatable(., options = list(pageLength = 10))
```

Now I want to short this list for those players who played 100 games.

```{r}
df100<- players_overall %>% 
  filter(nWin > 100)
  arrange(desc(nWin)) %>% 
  as.data.frame()
```

Now let's see the top 20 of the players

```{r}
top20 <- df100 %>% 
    arrange(desc(winPerc)) %>% 
    head(n = 20) %>%
    as.data.frame()

top20 %>% 
  select(Name, nWin, nLose, winPerc) %>% 
  datatable(., options = list(pageLength = 10))

```

#### About our top 20

According to our data, Federer is the player who won the most matches from 2000 through 2016 for a significant advantage. Followed by Nadal and Djokovic.

The next places at the top of the players behave with a constant gap because unlike Federer who has a bigger advantage, the next players in the table tend to have a smaller trend.

```{r}
top20 %>% 
  ggplot(aes(y = Name , x = nWin , fill = nWin )) +
  geom_bar(stat="identity",position=position_dodge(), alpha = 0.8) + theme_minimal() + 
  scale_fill_gradient(low="#4f908c",high="#6e0ff9") +  theme(legend.position="none")+
  geom_text(aes(label= nWin), hjust= -0.2)
```

#### About the surface. 

Now let's see if the type of field affects the performance of the players, to have an understanding of the field circumstances.

According to the heat map, Federer and Nadal have similar competition on the "Grass" and "Clays" fields, finding significant competitiveness between both players.

However, the circumstance of the surface is not something that determines whether or not the player will win the game.

```{r}
ggplot(data.frame(players_bySurface %>% 
                    filter (Name %in% players_overall[1:20, "Name"] & Surface != "Carpet")), aes(x = Surface, y = Name)) +
  geom_raster(aes(fill = winPerc))+
  labs(x = "Surface", y = "Name")+
  scale_fill_gradientn(colours=c("#cc0000","#eeeeee","#0000ff"))
```

## About the odds agencies.

Now we will talk about the Igaming agencies and how they have behaved over time, to know which one is best for me to invest and bet on.

I'm sure Federer will win me a couple of bucks! But let's see which agency tends to have a better prediction of games won and lost.

First, let me create a dataframe for this analysis.

```{r}
agencies = Data [,12:55]
rank_agencies <- select(agencies, -3:-15)
rank_agencies2 <- select(rank_agencies, -27:-30) 
rank_agencies2[3:20] <- apply(rank_agencies2[3:20], 2, round, 0)

rankagencies = rank_agencies2 [,1:27]
```

Now, I'm going to clean and manipulate some data to have a better understanding.

```{r}
#Let's create columns with the wins and loses of each company

#CB Win and lose odds
rankagencies$CBW_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$CBW, 1, 0)
rankagencies$CB_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$CBL, 1, 0)

#GB Win and lose odds
rankagencies$GB_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$GBW, 1, 0)
rankagencies$GB_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$GBL, 1, 0)

#IW Win and lose odds
rankagencies$IW_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$IWW, 1, 0)
rankagencies$IW_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$IWL, 1, 0)

#SB Win and lose odds
rankagencies$SB_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$SBW, 1, 0)
rankagencies$SB_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$SBL, 1, 0)

#B365 Win and lose odds
rankagencies$B365_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$B365W, 1, 0)
rankagencies$B365_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$B365L, 1, 0)

#EX Win and lose odds
rankagencies$EX_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$EXW, 1, 0)
rankagencies$EX_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$EXL, 1, 0)

#PS Win and lose odds
rankagencies$PS_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$PSW, 1, 0)
rankagencies$PS_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$PSL, 1, 0)

#UB Win and lose odds
rankagencies$UB_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$UBW, 1, 0)
rankagencies$UB_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$UBL, 1, 0)

#LB Win and lose odds
rankagencies$LB_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$LBW, 1, 0)
rankagencies$LB_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$LBL, 1, 0)

#SJ Win and lose odds
rankagencies$SJ_Win_Odds <- ifelse(rankagencies$WRank == rankagencies$SJW, 1, 0)
rankagencies$SJ_Lose_Odds <- ifelse(rankagencies$LRank == rankagencies$SJL, 1, 0)

```

Okay then, lets start analyzing these.

```{r}
summary(rankagencies)
scores = rankagencies [,27:47]

scores %>% 
  group_by(Season) %>%
  summarize(count = n())
```

```{r}
Win_Ods = scores %>% 
  select(Season, ends_with("Win_Odds")) 

Loser_Ods = scores %>% 
  select(Season, ends_with("Lose_Odds")) 

summary(Win_Ods)

TotalWins = Win_Ods %>%
  group_by(Season) %>%
  summarize_all(funs(sum(., na.rm = TRUE)))

TotalLose = TotalLose = Loser_Ods %>%
  group_by(Season) %>%
  summarize_all(funs(sum(., na.rm = TRUE)))

summary(Loser_Ods)

TotL = TotalLose %>% 
  summarize_all(funs(sum(., na.rm = TRUE)))
```

#### Win odds

According to the analysis, between 2000 and 20016, the first agency began betting in 2002. We can see a clear competition between the agencies, and not so much in others... In a moment we will do a more detailed analysis to see in Which one should I bet?

```{r}
Win_Ods_long <- melt(Win_Ods, id.vars = "Season")

ggplot(Win_Ods_long, aes(x = Season, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") +
  ggtitle("Win Odds by Year") +
  xlab("Year") +
  ylab("Win Odds") +
  theme(legend.title=element_blank())
```

#### Top 3 win odds

```{r}
Top3W = TotalWins %>% 
  select(Season, B365_Win_Odds,EX_Win_Odds,PS_Win_Odds)

Top3W_long <- melt(Top3W, id.vars = "Season")

ggplot(Top3W_long, aes(x = Season, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") +
  ggtitle("Win Odds by Year") +
  xlab("Year") +
  ylab("Win Odds") +
  theme(legend.title=element_blank())
```

Here we can see that the agency that has been betting for the longest time, since 2002, has been B365.

And in this top 3, in a total of games that they predicted and won, the results were as follows:

```{r}
Win_Ods %>%
  summarize_all(funs(sum(!is.na(.))))
```

```{r}
Winss = Win_Ods %>%
  summarize_all(funs(sum(!is.na(.))))

Winss$Season<-NULL

Winss_long <- Winss %>% 
  gather(key = "columns", value = "value") 

ggplot(Winss_long, aes(x = columns, y = value)) + geom_col(fill = "blue")+
  theme(axis.text.x = element_text(angle = 45))

```

#### 
